Prediction of PIK3CA mutation with gene expression

### 2020-05-12 plosOne/plosOne.Rmd
gene predictors

---
title: Prediction of PIK3CA mutation with gene expression
author:
  - name: Jun Kang
    email: jkang.alien@gmail.com
    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea
    corresponding: jkang.alien@gmail.com
abstract: |
bibliography: [mybibfile.bib, pik3ca.bib, rpackage.bib] 
indent: true
output: rticles::plos_article
csl: plos.csl
---

@@

Machine learning approach has been actively researched to detect the genomic alterations. Machine learning can build a prediction model from a large number of predictors such as radiolomic features @dercleIdentificationNonSmall2020, pathology image or gene expression data @wayMachineLearningDetects2018. Because most direct genomic test are more specific and sensitive than predictive models, machine learning approach might has limited role in clinical practice. However the machine learning prediction can be a second best when the direct test fails.

@@

PIK3CA is encodes the p110$\alpha$ catalytic subunit of phosphatidylinositol 3′-kinase (PI3K). PI3K is a protein kinase which phosphorylates phosphatidylinositol 4,5-biphosphate (PIP~2~) to make phosphatidylinositol 3,4,5-triphosphate (PIP~3~). Phosphatase and tensin homolog (PTEN) changes PIP~2~ to PIP~3~ in contrast PI3K. PIP~3~ is a second mesenger to activate protein kinase B (AKT) which is a serine/threonine-specific protein kinase. AKT inhibits apoptosis and promote cell proliferation.

@@

We used TCGA pancancer dataset. PIK3CA mutation data was get using cgdsr rpackage.[@jacobsenCgdsrRbasedAPI2019] Gene expression data was get from GDAC firehose using RTCGAToolbox R package. [@samurRTCGAToolboxNewTool2014]

@@

To narrow down potential predictors, Genes with a large the median absolute deviation (more than third-quartiles) were selected.  Yeo-Johnson transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package. [@kuhnRecipesPreprocessingTools2020] Penalized logistic regression was applied to prediction modeling. 10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}. Lambda is  penalty scaling parameter and alpha is mixing parameter of penalty function ($(1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1$). [@friedmanRegularizationPathsGeneralized2010] Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR). The modelin process was done with tidymodels rpackage.  [@kuhnTidymodelsEasilyInstall2020]

@@

## prevalance rate rate of PIK3CA mutation   
prevalance rate rate of PIK3CA was 0.11 in all cases. The PIK3CA prevalance rate of each cancer type was vary. The median prevalence of PIK3CA of each cancer types  was 0.03 (range 0-0.33) (Figure 1).

@@

In this study, the split method was done for testset instead of indepedent testset. Because there is no standard analysis method for the RNA-seq gene expression quantification, it is difficult to find gene expression data same as TGCA data. Developing standard analysis method for gene expression data is neccessory to apply this prediction model.

In this study, the prevalence of PIK3CA mutation was 11%. The low prevalence of PIK3CA results in imbalaced dataset which makes the prediction difficult. The AUROC is optimistic in terms of performance. The AUPR is more informative than AUROC on imbalaced datasets. [@saitoPrecisionRecallPlotMore2015]

@@

Our PIK3CA prediction model performed better than RAS activation prediction model of previous study in terms of both AUROC (0.84 vs 0.75) and AUPR (0.39 vs 0.24) on testset prediction. Our testset is corresponding to the samples initially filtered from training. The target variable of our study is more specific than the previous study. The specific important mutations can effect stronger downsteam gene expression than the broad events pathway activation. The previous study might be more difficult prediction problum.

@@

>>>>

---
title: Prediction of PIK3CA mutation with gene expression
author:
  - name: Jun Kang
    email: jkang.alien@gmail.com
    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea
    corresponding: jkang.alien@gmail.com
abstract: |
bibliography: [mybibfile.bib, pik3ca.bib, rpackage.bib, genes.bib] 
indent: true
output: rticles::plos_article
csl: plos.csl
---

@@

Machine learning approach has been actively researched to detect the genomic alterations. Machine learning can build a prediction model from a large number of predictors such as radiolomic features @dercleIdentificationNonSmall2020, pathology image [@coudrayClassificationMutationPrediction2018] or gene expression data [@wayMachineLearningDetects2018 @loiPIK3CAMutationsAssociated2010]. Because most direct genomic test are more specific and sensitive than predictive models, machine learning approach might has limited role in clinical practice. However the machine learning prediction can be a second best when the direct test fails.

@@

PIK3CA is encodes the p110$\alpha$ catalytic subunit of phosphatidylinositol 3′-kinase (PI3K). PI3K is a protein kinase which phosphorylates phosphatidylinositol 4,5-biphosphate (PIP~2~) to make phosphatidylinositol 3,4,5-triphosphate (PIP~3~). Phosphatase and tensin homolog (PTEN) changes PIP~2~ to PIP~3~ in contrast PI3K. PIP~3~ is a second mesenger to activate protein kinase B (AKT) which is a serine/threonine-specific protein kinase. AKT inhibits apoptosis and promote cell proliferation. @cantleyPhosphoinositide3KinasePathway2002

@@

We used TCGA pancancer dataset. TCGA is a cancer genomic consotium that archives data of exom sequencing, gene expression, DNA methylation, protein expression and clicial data of more than 10000 cancer samples across 33 common cancer types. The gene expression TCGA pancancer dataset is batch-corrected with normalization. TCGA dataset is publically available. PIK3CA mutation data was get using cgdsr rpackage.[@jacobsenCgdsrRbasedAPI2019] Gene expression data was get from GDAC firehose using RTCGAToolbox R package. [@samurRTCGAToolboxNewTool2014]

@@

To narrow down potential predictors, Genes with a large the median absolute deviation (more than third-quartiles) were selected.  Yeo-Johnson transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package. [@kuhnRecipesPreprocessingTools2020] Penalized logistic regression was applied to prediction modeling. 10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}. Lambda is  penalty scaling parameter and alpha is mixing parameter of penalty function ($(1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1$). [@friedmanRegularizationPathsGeneralized2010]

## Accessing model performance

Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR). The prevalence of PIK3CA mutation is low. The low prevalence of PIK3CA results in imbalaced dataset which makes the prediction difficult. The AUROC is optimistic in terms of performance. The AUPR is more informative than AUROC on imbalaced datasets. [@saitoPrecisionRecallPlotMore2015]   The modeling process was done with tidymodels rpackage.  [@kuhnTidymodelsEasilyInstall2020]

@@

## prevalance rate rate of PIK3CA mutation   
Prevalance rate of PIK3CA was 0.11 in all cases. The PIK3CA prevalance rate of each cancer type was vary. The median prevalance rate of PIK3CA of each cancer types  was 0.03 (range 0-0.33) (Figure 1).

@@

Prediction modeling from TCGA pancancer dataset has limitations regarding to data preprocessing. Methods for gancer gene expression has been developed for research. To stastical analysis, the gene expression data is processed between-sample normalization to remove batch effect. If the model has been trained from the between-sample normalization, a new sample can not be exactly processed like trainset. A model based on gene expression TCGA pancacner dataset has limitation on preprocessing. It is nessessory developing preprocessing method which is independent with dataset to apply the gene expression data to prediction model.

@@

Our PIK3CA prediction model performed better than RAS activation prediction model of previous study in terms of both AUROC (0.84 vs 0.75) and AUPR (0.39 vs 0.24) on testset prediction. Our testset is corresponding to the samples initially filtered from training. The target variable of our study is more specific than the previous study. The specific important mutations can effect stronger downsteam gene expression than the broad events pathway activation. The previous study might be more difficult prediction problem.

@@

Some significant gene expression predictors were closely related with PTEN and the PI3K pathway. PTEN and IGFR1R are the strongest gene expression predictor which have negative and positive predictive power. IGF1R is a tyrosine kinase receptor which activates PI3K. [@leroithInsulinlikeGrowthFactor2003] Insulin receptor substrate-2 (IRS2) is the adaptor protein of IGF1R. [@heInteractionInsulinReceptor1996] PTEN is an important regulator of PIP~3~ by dephosphorylating  PIP~3~ in constrast PI3K.[@cantleyPhosphoinositide3KinasePathway2002]

Another study of PIK3CA mutation prediction showed good performance AUROC 0.71 in independent testset. They made gene-expression signature which is sum of the average of the logarithmic gene expression. [@cizkovaGeneExpressionProfiling2010; @loiPIK3CAMutationsAssociated2010]

Another study predicted copy numbear alterations with gene expression using multinomial logistic regression model with least absolute shrinkage and selection operator (LASSO). The prediction of 1p/19q codel was very good with an AUROC of 0.997. The gene level prediction was good with an AUROC 0.75. [@muCNAPEMachineLearning2019]

A Hidden Markov Model Approach for Prediction of Genomic Alterations from Gene Expression Profiling.[@gengHiddenMarkovModel2008]

[@heGeneSignaturesAssociated2020]

### 2020-05-12 plosOne/plosOne.tex
gene predictors

Targeted thrapy becomes standard treatment in many cancer patients. To
treat with the targeted therapy, the The patients should be confirmed
the indicated genomic alterations in their cancer. Many direct test for
the genomic alteration has been developed and prooved for their clical
utility to discriminate which patients will be response the targeted
therapy.

The machine learning approach has been actively researched recently to
detect the genomic alterations. The machine learning can build a
prediction model from a large number of predictors such as radiolomic
features {[}1{]}, pathology image or gene expression data {[}2{]}. The
machine learning approach has limitations to apply to clinical practice.
However the machine learning prediction can be a second best when the
direct test fails.

PIK3CA is encodes the p110\(\alpha\) catalytic subunit of
phosphatidylinositol 3′-kinase (PI3K). PI3K is a protein kinase which
phosphorylates phosphatidylinositol 4,5-biphosphate
(PIP\textsubscript{2}) to make phosphatidylinositol 3,4,5-triphosphate
(PIP\textsubscript{3}). Phosphatase and tensin homolog (PTEN) changes
PIP\textsubscript{2} to PIP\textsubscript{3} in contrast PI3K.
PIP\textsubscript{3} is a second mesenger to activate protein kinase B
(AKT) which is a serine/threonine-specific protein kinase. AKT inhibits
apoptosis and promote cell proliferation.

Breast cancer with PIK3CA mutation has been approved to use PIK3CA
inhibitor in hormone receptor positive HER2 negative subtype. {[}3{]}
The PIK3CA mutation is second most driver mutation after TP53. The
PIK3CA mutation is most frequantly founded in endometrial carcinoma
(45\%), and followed by breast invasive carcinoma (24\%), cervical
squamous cell carcinoma and endocervical adenocarcinoma (20\%) and colon
adenocarcinoma (16\%).

Prediction RAS pathway activation with gene expression data was done in
previous study. {[}4{]} They trained pancancer The Cancer Genome Atlas
(TCGA) data with a supervised elastic net penalized logistic regression
classifier with stochastic gradient descent. The performance of their
model was 84\% with an area under the receiver operating characteristic
(AUROC) curve and 63\% with an area under the precision recall (AUPR)
curve. We applied their modeling methods in prediction PIK3CA mutation.

@@

We used TCGA pancancer dataset. PIK3CA mutation data was get using cgdsr
rpackage.{[}5{]} Gene expression data was get from GDAC firehose using
RTCGAToolbox R package. {[}6{]}

10845 cases were available both PIK3CA mutation and mRNA expression
data. 5128 out of 20502 genes were included in the modeling process
after filtering with median absolute deviation as described at modelinf
process method. 33 cancer type dummy variables were included in
predictor variables.

The target variable was PIK3CA mutation status. The status of PIK3CA was
considered as positive when the case has following PIK3CA variants which
is the target variables of the therascreen PIK3CA RGQ PCR Kit; C420R,
E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
The therascreen PIK3CA RGQ PCR Kit was approved by U.S. Food and Drug
adminidtration for PIK3CA inhibitor. We splited the three quarters of
dataset for the trainset and one quarter for testset.

@@

To narrow down potential predictors, Genes with a large the median
absolute deviation (more than third-quartiles) were selected.
Yeo-Johnson transformation was done to correct skewness. Centering and
scaling were done. All preprocessing was done using recipe r package.
{[}7{]} Penalized logistic regression was applied to prediction
modeling. 10-fold cross-validation with targe variable stratification
was done over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
\(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Lambda is penalty scaling parameter
and alpha is mixing parameter of penalty function
(\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\)).
{[}8{]} Model performance was evealuated with the area under receiver
operating characteristic (ROC) and area under the ROC (AUROC) and
precision recall (PR) and area under curve (AUPR). The modelin process
was done with tidymodels rpackage. {[}9{]}

@@

\hypertarget{dataset-summary}{%
\subsection{Dataset summary}\label{dataset-summary}}

Prevalcence of PIK3CA was 0.11 in all cases. The PIK3CA prevalance of
each cancer type was vary. The median prevalence of PIK3CA of each
cancer types was 0.03 (range 0-0.33) (Figure 1).

@@

In 10-fold cross-validation, the model with lambda = 0.01 and alpha =
1.0 (Ridge regression) were best performance in terms of AUROC. The
final model was trained with the selected hyperparameters with all
trainset.

@@

Our model showed good performance to predict PIK3CA mutation of various
cancer types.

\begin{itemize}
\item
  Main message answers the question and main supporting evidence
\item
  Critical assessment opinions on any shortcomings in study design
\end{itemize}

In this study, the split method was done for testset instead of
indepedent testset. Because there is no standard analysis method for the
RNA-seq gene expression quantification, it is difficult to find gene
expression data same as TGCA data. Developing standard analysis method
for gene expression data is neccessory to apply this prediction model.

In this study, the prevalence of PIK3CA mutation was 11\%. The low
prevalence of PIK3CA results in imbalaced dataset which makes the
prediction difficult. The AUROC is optimistic in terms of performance.
The AUPR is more informative than AUROC on imbalaced datasets. {[}10{]}

@@

@@

\begin{itemize}
\item
  Figure 1. Prevalance rate of PIK3CA across cancer types The
  abbreviations of cancer types are explained in S1 appendix.
\item
  Figure 2. Summary of modeling results
\end{itemize}

@@

\leavevmode\hypertarget{ref-wayMachineLearningDetects2018}{}%
2. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
Genome Atlas. Cell Reports. Elsevier; 2018;23: 172--180.e3.
doi:\href{https://doi.org/10.1016/j.celrep.2018.03.046}{10.1016/j.celrep.2018.03.046}

\leavevmode\hypertarget{ref-andre_2019_alpelisib_nengljmeda}{}%
3. André F, Ciruelos E, Rubovszky G, Campone M, Loibl S, Rugo HS, et al.
Alpelisib for PIK3CA-Mutated, Hormone ReceptorPositive Advanced Breast
Cancer. New England Journal of Medicine. Massachusetts Medical Society;
2019;380: 1929--1940.
doi:\href{https://doi.org/10.1056/NEJMoa1813904}{10.1056/NEJMoa1813904}

\leavevmode\hypertarget{ref-way_2018_machine_cellreports}{}%
4. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
Genome Atlas. Cell Reports. Elsevier; 2018;23: 172--180.e3.
doi:\href{https://doi.org/10.1016/j.celrep.2018.03.046}{10.1016/j.celrep.2018.03.046}

\leavevmode\hypertarget{ref-jacobsenCgdsrRbasedAPI2019}{}%
5. Jacobsen A, Luna A. Cgdsr: R-based API for accessing the MSKCC cancer
genomics data server (CGDS). 2019.

\leavevmode\hypertarget{ref-samurRTCGAToolboxNewTool2014}{}%
6. Samur MK. RTCGAToolbox: A new tool for exporting TCGA Firehose data.
PLoS One. 2014;9(9):e106397.

\leavevmode\hypertarget{ref-kuhnRecipesPreprocessingTools2020}{}%
7. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
matrices. 2020.

\leavevmode\hypertarget{ref-friedmanRegularizationPathsGeneralized2010}{}%
8. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
Generalized Linear Models via Coordinate Descent. J Stat Softw. 2010;33:
1--22.

\leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
9. Kuhn M, Wickham H. Tidymodels: Easily install and load the
'tidymodels' packages. 2020.

\leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
10. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
than the ROC Plot When Evaluating Binary Classifiers on Imbalanced
Datasets. PLOS ONE. Public Library of Science; 10: e0118432.
doi:\href{https://doi.org/10.1371/journal.pone.0118432}{10.1371/journal.pone.0118432}

>>>>

Targeted thrapy becomes standard treatment in many cancer patients. Many
targeted therapy requires test for a specific cancer genomic alteration,
to treat the patients. Many direct test for the genomic alteration has
been developed and prooved for their clical utility to discriminate
which patients will be response the targeted therapy.

Machine learning approach has been actively researched to detect the
genomic alterations. Machine learning can build a prediction model from
a large number of predictors such as radiolomic features {[}1{]},
pathology image {[}2{]} or gene expression data {[}3{]}. Because most
direct genomic test are more specific and sensitive than predictive
models, machine learning approach might has limited role in clinical
practice. However the machine learning prediction can be a second best
when the direct test fails.

Prediction RAS pathway activation with gene expression data was done in
previous study. {[}5{]} They trained pancancer The Cancer Genome Atlas
(TCGA) data with a supervised elastic net penalized logistic regression
classifier with stochastic gradient descent. The performance of their
model was 84\% with an area under the receiver operating characteristic
(AUROC) curve and 63\% with an area under the precision recall (AUPR)
curve. The authors suggested their approach can be applied to other
genomic alterations.

PIK3CA is encodes the p110\(\alpha\) catalytic subunit of
phosphatidylinositol 3′-kinase (PI3K). PI3K is a protein kinase which
phosphorylates phosphatidylinositol 4,5-biphosphate
(PIP\textsubscript{2}) to make phosphatidylinositol 3,4,5-triphosphate
(PIP\textsubscript{3}). Phosphatase and tensin homolog (PTEN) changes
PIP\textsubscript{2} to PIP\textsubscript{3} in contrast PI3K.
PIP\textsubscript{3} is a second mesenger to activate protein kinase B
(AKT) which is a serine/threonine-specific protein kinase. AKT inhibits
apoptosis and promote cell proliferation. {[}6{]}

Breast cancer with PIK3CA mutation has been approved to use PIK3CA
inhibitor in hormone receptor positive HER2 negative subtype. {[}7{]}
The PIK3CA mutation is second most driver mutation after TP53. The
PIK3CA mutation is most frequantly founded in endometrial carcinoma
(45\%), and followed by breast invasive carcinoma (24\%), cervical
squamous cell carcinoma and endocervical adenocarcinoma (20\%) and colon
adenocarcinoma (16\%).

We apply a supervised elastic net penalized logistic regression model in
prediction PIK3CA mutation. The purpose of this study is to know this
prediction model approach can be applied not only RAS pathway activation
but also PIK3CA mutation across many cancer types.

@@

We used TCGA pancancer dataset. TCGA is a cancer genomic consotium that
archives data of exom sequencing, gene expression, DNA methylation,
protein expression and clicial data of more than 10000 cancer samples
across 33 common cancer types. The gene expression TCGA pancancer
dataset is batch-corrected with normalization. TCGA dataset is
publically available. PIK3CA mutation data was get using cgdsr
rpackage.{[}8{]} Gene expression data was get from GDAC firehose using
RTCGAToolbox R package. {[}9{]}

10845 cases were available both PIK3CA mutation and mRNA expression
data. 5128 out of 20502 genes were included in the modeling process
after filtering with median absolute deviation as described at modeling
process method. 33 cancer type dummy variables were included in
predictor variables.

The target variable was PIK3CA mutation status. The status of PIK3CA was
considered as positive when the case has following PIK3CA variants which
is the target variables of the therascreen PIK3CA RGQ PCR Kit; C420R,
E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
The therascreen PIK3CA RGQ PCR Kit was approved as a companion diagnosis
to treat with PIK3CA inhibitor by U.S. Food and Drug adminidtration. We
splited the three quarters of dataset for the trainset and one quarter
for testset.

@@

To narrow down potential predictors, Genes with a large the median
absolute deviation (more than third-quartiles) were selected.
Yeo-Johnson transformation was done to correct skewness. Centering and
scaling were done. All preprocessing was done using recipe r package.
{[}10{]} Penalized logistic regression was applied to prediction
modeling. 10-fold cross-validation with targe variable stratification
was done over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
\(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Lambda is penalty scaling parameter
and alpha is mixing parameter of penalty function
(\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\)).
{[}11{]}

\hypertarget{accessing-model-performance}{%
\subsection{Accessing model
performance}\label{accessing-model-performance}}

Model performance was evealuated with the area under receiver operating
characteristic (ROC) and area under the ROC (AUROC) and precision recall
(PR) and area under curve (AUPR). The prevalence of PIK3CA mutation is
low. The low prevalence of PIK3CA results in imbalaced dataset which
makes the prediction difficult. The AUROC is optimistic in terms of
performance. The AUPR is more informative than AUROC on imbalaced
datasets. {[}12{]} The modeling process was done with tidymodels
rpackage. {[}13{]}

@@

\hypertarget{prevalance-rate-rate-of-pik3ca-mutation}{%
\subsection{prevalance rate rate of PIK3CA
mutation}\label{prevalance-rate-rate-of-pik3ca-mutation}}

Prevalance rate of PIK3CA was 0.11 in all cases. The PIK3CA prevalance
rate of each cancer type was vary. The median prevalance rate of PIK3CA
of each cancer types was 0.03 (range 0-0.33) (Figure 1).

@@

In 10-fold cross-validation, the model with \(\lambda\) = 0.01 and
\(\alpha\) = 1.0 (Ridge regression) showed best performance in terms of
AUROC. The final model was trained with the selected hyperparameters
with all trainset.

@@

\begin{itemize}
\tightlist
\item
  Main message answers the question and main supporting evidence
\end{itemize}

Our model showed good performance to predict PIK3CA mutation of various
cancer types. This result shows that the supervised elastic net
penalized logistic regression model can be applied not only RAS
activation pathway but also PIK3CA mutation. Both RAS activation pathway
and PIK3CA mutation are important and common cancer genomic alterations.
They have significant effect on gene expression in cacer cells. It might
be challenging prediction of genomic alterations which are infrequent or
have weak effect on gene expression.

\begin{itemize}
\tightlist
\item
  Critical assessment opinions on any shortcomings in study design
\end{itemize}

Prediction modeling from TCGA pancancer dataset has limitations
regarding to data preprocessing. Methods for gancer gene expression has
been developed for research. To stastical analysis, the gene expression
data is processed between-sample normalization to remove batch effect.
If the model has been trained from the between-sample normalization, a
new sample can not be exactly processed like trainset. A model based on
gene expression TCGA pancacner dataset has limitation on preprocessing.
It is nessessory developing preprocessing method which is independent
with dataset to apply the gene expression data to prediction model.

@@

Some significant gene expression predictors were closely related with
PTEN and the PI3K pathway. PTEN and IGFR1R are the strongest gene
expression predictor which have negative and positive predictive power.
IGF1R is a tyrosine kinase receptor which activates PI3K. {[}14{]}
Insulin receptor substrate-2 (IRS2) is the adaptor protein of IGF1R.
{[}15{]} PTEN is an important regulator of PIP\textsubscript{3} by
dephosphorylating PIP\textsubscript{3} in constrast PI3K.{[}6{]}

Another study of PIK3CA mutation prediction showed good performance
AUROC 0.71 in independent testset. They made gene-expression signature
which is sum of the average of the logarithmic gene expression.
{[}4,16{]}

Another study predicted copy numbear alterations with gene expression
using multinomial logistic regression model with least absolute
shrinkage and selection operator (LASSO). The prediction of 1p/19q codel
was very good with an AUROC of 0.997. The gene level prediction was good
with an AUROC 0.75. {[}17{]}

A Hidden Markov Model Approach for Prediction of Genomic Alterations
from Gene Expression Profiling.{[}18{]}

{[}19{]}

@@

\begin{itemize}
\item
  Figure 1. prevalance rate rate of PIK3CA across cancer types The
  abbreviations of cancer types are explained in S1 appendix.
\item
  Figure 2. Summary of modeling results
\end{itemize}

@@

\leavevmode\hypertarget{ref-coudrayClassificationMutationPrediction2018}{}%
2. Coudray N, Ocampo PS, Sakellaropoulos T, Narula N, Snuderl M, Fenyö
D, et al. Classification and mutation prediction from nonSmall cell lung
cancer histopathology images using deep learning. Nature Medicine.
Nature Publishing Group; 2018;24: 1559--1567.
doi:\href{https://doi.org/10.1038/s41591-018-0177-5}{10.1038/s41591-018-0177-5}

\leavevmode\hypertarget{ref-wayMachineLearningDetects2018}{}%
3. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
Genome Atlas. Cell Reports. Elsevier; 2018;23: 172--180.e3.
doi:\href{https://doi.org/10.1016/j.celrep.2018.03.046}{10.1016/j.celrep.2018.03.046}

\leavevmode\hypertarget{ref-loiPIK3CAMutationsAssociated2010}{}%
4. Loi S, Haibe-Kains B, Majjaj S, Lallemand F, Durbecq V, Larsimont D,
et al. PIK3CA mutations associated with gene signature of low mTORC1
signaling and better outcomes in estrogen receptorPositive breast
cancer. PNAS. National Academy of Sciences; 2010;107: 10208--10213.
doi:\href{https://doi.org/10.1073/pnas.0907011107}{10.1073/pnas.0907011107}

\leavevmode\hypertarget{ref-way_2018_machine_cellreports}{}%
5. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
Genome Atlas. Cell Reports. Elsevier; 2018;23: 172--180.e3.
doi:\href{https://doi.org/10.1016/j.celrep.2018.03.046}{10.1016/j.celrep.2018.03.046}

\leavevmode\hypertarget{ref-cantleyPhosphoinositide3KinasePathway2002}{}%
6. Cantley LC. The Phosphoinositide 3-Kinase Pathway. Science. American
Association for the Advancement of Science; 2002;296: 1655--1657.
doi:\href{https://doi.org/10.1126/science.296.5573.1655}{10.1126/science.296.5573.1655}

\leavevmode\hypertarget{ref-andre_2019_alpelisib_nengljmeda}{}%
7. André F, Ciruelos E, Rubovszky G, Campone M, Loibl S, Rugo HS, et al.
Alpelisib for PIK3CA-Mutated, Hormone ReceptorPositive Advanced Breast
Cancer. New England Journal of Medicine. Massachusetts Medical Society;
2019;380: 1929--1940.
doi:\href{https://doi.org/10.1056/NEJMoa1813904}{10.1056/NEJMoa1813904}

\leavevmode\hypertarget{ref-jacobsenCgdsrRbasedAPI2019}{}%
8. Jacobsen A, Luna A. Cgdsr: R-based API for accessing the MSKCC cancer
genomics data server (CGDS). 2019.

\leavevmode\hypertarget{ref-samurRTCGAToolboxNewTool2014}{}%
9. Samur MK. RTCGAToolbox: A new tool for exporting TCGA Firehose data.
PLoS One. 2014;9(9):e106397.

\leavevmode\hypertarget{ref-kuhnRecipesPreprocessingTools2020}{}%
10. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
matrices. 2020.

\leavevmode\hypertarget{ref-friedmanRegularizationPathsGeneralized2010}{}%
11. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
Generalized Linear Models via Coordinate Descent. J Stat Softw. 2010;33:
1--22.

\leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
12. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
than the ROC Plot When Evaluating Binary Classifiers on Imbalanced
Datasets. PLOS ONE. Public Library of Science; 10: e0118432.
doi:\href{https://doi.org/10.1371/journal.pone.0118432}{10.1371/journal.pone.0118432}

\leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
13. Kuhn M, Wickham H. Tidymodels: Easily install and load the
'tidymodels' packages. 2020.

\leavevmode\hypertarget{ref-leroithInsulinlikeGrowthFactor2003}{}%
14. LeRoith D, Roberts CT. The insulin-like growth factor system and
cancer. Cancer Letters. 2003;195: 127--137.
doi:\href{https://doi.org/10.1016/S0304-3835(03)00159-9}{10.1016/S0304-3835(03)00159-9}

\leavevmode\hypertarget{ref-heInteractionInsulinReceptor1996}{}%
15. He W, Craparo A, Zhu Y, O'Neill TJ, Wang L-M, Pierce JH, et al.
Interaction of Insulin Receptor Substrate-2 (IRS-2) with the Insulin and
Insulin-like Growth Factor I Receptors EVIDENCE FOR TWO DISTINCT
PHOSPHOTYROSINE-DEPENDENT INTERACTION DOMAINS WITHIN IRS-2. J Biol Chem.
American Society for Biochemistry and Molecular Biology; 1996;271:
11641--11645.
doi:\href{https://doi.org/10.1074/jbc.271.20.11641}{10.1074/jbc.271.20.11641}

\leavevmode\hypertarget{ref-cizkovaGeneExpressionProfiling2010}{}%
16. Cizkova M, Cizeron-Clairac G, Vacher S, Susini A, Andrieu C,
Lidereau R, et al. Gene Expression Profiling Reveals New Aspects of
PIK3CA Mutation in ERalpha-Positive Breast Cancer: Major Implication of
the Wnt Signaling Pathway. PLOS ONE. Public Library of Science; 5:
e15647.
doi:\href{https://doi.org/10.1371/journal.pone.0015647}{10.1371/journal.pone.0015647}

\leavevmode\hypertarget{ref-muCNAPEMachineLearning2019}{}%
17. Mu Q, Wang J. CNAPE: A Machine Learning Method for Copy Number
Alteration Prediction from Gene Expression. IEEE/ACM Transactions on
Computational Biology and Bioinformatics. 2019; 1--1.
doi:\href{https://doi.org/10.1109/TCBB.2019.2944827}{10.1109/TCBB.2019.2944827}

\leavevmode\hypertarget{ref-gengHiddenMarkovModel2008}{}%
18. Geng H, Ali HH, Chan WC. A Hidden Markov Model Approach for
Prediction of Genomic Alterations from Gene Expression Profiling. In:
Măndoiu I, Sunderraman R, Zelikovsky A, editors. Bioinformatics Research
and Applications. Berlin, Heidelberg: Springer; 2008. pp. 414--425.
doi:\href{https://doi.org/10.1007/978-3-540-79450-9_38}{10.1007/978-3-540-79450-9\_38}

\leavevmode\hypertarget{ref-heGeneSignaturesAssociated2020}{}%
19. He X, Qin C, Zhao Y, Zou L, Zhao H, Cheng C. Gene signatures
associated with genomic aberrations predict prognosis in neuroblastoma.
Cancer Communications. 2020;40: 105--118.
doi:\href{https://doi.org/10.1002/cac2.12016}{10.1002/cac2.12016}
